\documentclass[aspectratio=169]{beamer}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{tikz}
\usepackage{fancyvrb}
\usepackage{xcolor}
\title{Symbolic Execution with Angr\\RPISEC}
\date{December 6, 2019}
\author{Avi Weinstock (\Verb|aweinstock|), Luke Biery (\Verb|tiecoon|)}

\input{../rpisec_beamer_theme.tex}

\begin{document}
\maketitle

\begin{frame}[fragile]
\frametitle{Overview}
\begin{itemize}
\item What is Symbolic Execution? What techniques does it compete with?
\item How symbolic execution works (theory)
\item How symbolic execution works (Angr commands)
\item Solving MBE lab1A with Angr
\end{itemize}
\end{frame}

\section{Background - What it is and what is the problem space?}
\begin{frame}[fragile]
\frametitle{What is Symbolic Execution?}
\begin{itemize}
\item Executes a program with symbolic data (usually input)
\item Essentially runs a program on "all possible inputs" at once
\item Instead of having concrete data in each variable/address, \\variables/addresses store trees of what to do with the input
\end{itemize}
\end{frame}

\begin{frame}[fragile]
\frametitle{What problems does Symbolic Execution solve?}
\begin{itemize}
\item What input to provide to reach/avoid a specific line of code?
\item How is a value deep in the program affected by some specific input?
\item Do any inputs lead to any crash?
\item On a crashing input, what registers are controlled by the input?
\end{itemize}
\end{frame}

\begin{frame}[fragile]
\frametitle{Symbolic Execution vs Fuzzing}
\begin{tabular}{c|c}
Symbolic Execution & Fuzzing\\\hline
\textcolor{green}{+ Explores all inputs} & \textcolor{red}{- Only explores random inputs}\\
\textcolor{red}{- Uses more memory/time} & \textcolor{green}{+ Uses around as much memory/time as target program}
\end{tabular}
TODO: more comparisons/columns? emphasize that "all inputs" means that symexec can find constant-time comparisons against a giant constant, unlike coverage-guided?
\end{frame}

\section{How symbolic execution works in general}
\begin{frame}[fragile]
\frametitle{TODO: Avi}
\begin{itemize}
\item symbolic branches
\item symbolic loops
\item symbolizing tainted memory (e.g. \verb|sym_memory[(EBP+username+i)]| vs \verb|sym_username[i]|)
\item state explosion when looping on symbolic data
\end{itemize}
\end{frame}

\section{How to use Angr for symbolic execution}
\begin{frame}[fragile]
\frametitle{TODO: Luke}
\begin{itemize}
\item loading binariess
\item marking input as symbolic
\item initiating the search/pruning the search space
\item simprocedures for shortcutting syscalls?
\end{itemize}
\end{frame}

\section{Example: MBE lab1A with Angr}
\begin{frame}[fragile]
\frametitle{Resources}
\begin{itemize}
\item \verb|https://github.com/angr/|
\item \verb|https://github.com/Z3Prover/z3/|
\item \verb|https://github.com/RPISEC/MBE|
\end{itemize}
\end{frame}
\end{document}
