\documentclass{beamer}
\usepackage{tikz}
\usepackage{fancyvrb}
\title{A Modular Integration of SAT/SMT Solvers to Coq through Proof Witnesses}
\date{April 24th 2018}
\author{
Paper: 
Michael Armand, Germain Faure, Benjamin Gregoire, Chantal Keller, Laurent Thery, Benjamin Werner\\
Presentation: Avi Weinstock
}
\begin{document}
\input{coqdoc_macros.tex}
\maketitle

%\section{Introduction}
%\frame{\tableofcontents[sectionstyle=show/shaded, subsectionstyle=show/shaded]}
\begin{frame}[fragile]
\frametitle{Gallina term syntax}
\begin{centerframe}\tiny
\begin{tabular}{lcl@{\quad~}r}  % warning: page width exceeded with \qquad 
{\term} & ::= &
         {\tt forall} {\binders} {\tt ,} {\term}  &\\
 & $|$ & {\tt fun} {\binders} {\tt =>} {\term} &\\
 & $|$ & {\tt fix} {\fixpointbodies} &\\
 & $|$ & {\tt cofix} {\cofixpointbodies} &\\
 & $|$ & {\tt let} {\ident} \zeroone{\binders} {\typecstr} {\tt :=} {\term}
         {\tt in} {\term} &\\
 & $|$ & {\tt let fix} {\fixpointbody} {\tt in} {\term} &\\
 & $|$ & {\tt let cofix} {\cofixpointbody}
         {\tt in} {\term} &\\
 & $|$ & {\tt let} {\tt (} \sequence{\name}{,} {\tt )} \zeroone{\ifitem}
         {\tt :=} {\term}
         {\tt in} {\term}  &\\
 & $|$ & {\tt let '} {\pattern} \zeroone{{\tt in} {\term}} {\tt :=} {\term}
        \zeroone{\returntype} {\tt in} {\term} & \\
 & $|$ & {\tt if} {\term} \zeroone{\ifitem} {\tt then} {\term}
         {\tt else} {\term} &\\
 & $|$ & {\term} {\tt :} {\term} &\\
 & $|$ & {\term} {\tt <:} {\term} &\\
 & $|$ & {\term} {\tt :>} &\\
 & $|$ & {\term} {\tt ->} {\term} &\\
 & $|$ & {\term} \nelist{\termarg}{}&\\
 & $|$ & {\tt @} {\qualid} \sequence{\term}{}
            &\\
 & $|$ & {\term} {\tt \%} {\ident} &\\
 & $|$ & {\tt match} \nelist{\caseitem}{\tt ,}
                 \zeroone{\returntype} {\tt with} &\\
    &&   ~~~\zeroone{\zeroone{\tt |} \nelist{\eqn}{|}} {\tt end}
    &\\
 & $|$ & {\qualid} &\\
 & $|$ & {\sort} &\\
 & $|$ & {\num} &\\
 & $|$ & {\_} &\\
 & $|$ & {\tt (} {\term} {\tt )} & \\
 & & &\\
\end{tabular}
\end{centerframe}
\end{frame}

\begin{frame}[fragile]
\frametitle{Gallina term syntax}
\begin{centerframe}\tiny
\begin{tabular}{lcl@{\quad~}r}  % warning: page width exceeded with \qquad 
{\termarg} & ::= & {\term} &\\
 & $|$ & {\tt (} {\ident} {\tt :=} {\term} {\tt )}
         &\\
%% & $|$ & {\tt (} {\num} {\tt :=} {\term} {\tt )}
%%         &\\
&&&\\
{\binders} & ::= & \nelist{\binder}{}  \\
&&&\\
{\binder} & ::= &   {\name} &  \\
 & $|$ & {\tt (} \nelist{\name}{} {\tt :} {\term} {\tt )} &\\
 & $|$ & {\tt (} {\name} {\typecstr} {\tt :=} {\term} {\tt )} &\\
 & $|$ & {\tt '} {\pattern} &\\
& & &\\
{\name} & ::= & {\ident} &\\
 & $|$ & {\tt \_} &\\
&&&\\
{\qualid} & ::= & {\ident} & \\
 & $|$ & {\qualid} {\accessident} &\\
 & & &\\
{\sort} & ::= & {\tt Prop} ~$|$~ {\tt Set} ~$|$~ {\tt Type} &
\end{tabular}
\end{centerframe}
\end{frame}

\begin{frame}[fragile]
\frametitle{CIC Derivation Trees}
\begin{description}
\item[W-Empty] \inference{\WF{[]}{}}
\item[W-Local-Assum]  % Ce n'est pas vrai : x peut apparaitre plusieurs fois dans Gamma
\inference{\frac{\WTEG{T}{s}~~~~s \in \Sort~~~~x \not\in \Gamma % \cup E
      }{\WFE{\Gamma::(x:T)}}}
\item[W-Local-Def]
\inference{\frac{\WTEG{t}{T}~~~~x \not\in \Gamma % \cup E
     }{\WFE{\Gamma::(x:=t:T)}}}
\item[W-Global-Assum] \inference{\frac{\WTE{}{T}{s}~~~~s \in \Sort~~~~c \notin E}
                      {\WF{E;c:T}{}}}
\item[W-Global-Def] \inference{\frac{\WTE{}{t}{T}~~~c \notin E}
                      {\WF{E;c:=t:T}{}}}
\item[Ax-Prop] \index{Typing rules!Ax-Prop}
\inference{\frac{\WFE{\Gamma}}{\WTEG{\Prop}{\Type(1)}}}
\end{description}
\end{frame}

\begin{frame}[fragile]
\frametitle{CIC Derivation Trees}
\begin{description}
\item[Ax-Set] \index{Typing rules!Ax-Set}
\inference{\frac{\WFE{\Gamma}}{\WTEG{\Set}{\Type(1)}}}
\item[Ax-Type] \index{Typing rules!Ax-Type}
\inference{\frac{\WFE{\Gamma}}{\WTEG{\Type(i)}{\Type(i+1)}}}
\item[Var]\index{Typing rules!Var}
 \inference{\frac{ \WFE{\Gamma}~~~~~(x:T) \in \Gamma~~\mbox{or}~~(x:=t:T) \in \Gamma~\mbox{for some $t$}}{\WTEG{x}{T}}}
\item[Const]  \index{Typing rules!Const}
\inference{\frac{\WFE{\Gamma}~~~~(c:T) \in E~~\mbox{or}~~(c:=t:T) \in E~\mbox{for some $t$} }{\WTEG{c}{T}}}
\end{description}
\end{frame}

\begin{frame}[fragile]
\frametitle{CIC Derivation Trees}
\begin{description}
\item[Prod-Prop]  \index{Typing rules!Prod-Prop}
\inference{\frac{\WTEG{T}{s}~~~~s \in \Sort~~~
    \WTE{\Gamma::(x:T)}{U}{\Prop}}
      { \WTEG{\forall~x:T,U}{\Prop}}} 
\item[Prod-Set]  \index{Typing rules!Prod-Set}
\inference{\frac{\WTEG{T}{s}~~~~s \in\{\Prop, \Set\}~~~~~~
    \WTE{\Gamma::(x:T)}{U}{\Set}}
      { \WTEG{\forall~x:T,U}{\Set}}} 
\item[Prod-Type]  \index{Typing rules!Prod-Type}
\inference{\frac{\WTEG{T}{\Type(i)}~~~~
    \WTE{\Gamma::(x:T)}{U}{\Type(i)}}
    {\WTEG{\forall~x:T,U}{\Type(i)}}}
\end{description}
\end{frame}

\begin{frame}[fragile]
\frametitle{CIC Derivation Trees}
\begin{description}
\item[Lam]\index{Typing rules!Lam} 
\inference{\frac{\WTEG{\forall~x:T,U}{s}~~~~ \WTE{\Gamma::(x:T)}{t}{U}}
        {\WTEG{\lb x:T\mto t}{\forall x:T, U}}}
\item[App]\index{Typing rules!App}
 \inference{\frac{\WTEG{t}{\forall~x:U,T}~~~~\WTEG{u}{U}}
                 {\WTEG{(t\ u)}{\subst{T}{x}{u}}}}
\item[Let]\index{Typing rules!Let} 
\inference{\frac{\WTEG{t}{T}~~~~ \WTE{\Gamma::(x:=t:T)}{u}{U}}
        {\WTEG{\letin{x}{t:T}{u}}{\subst{U}{x}{t}}}}
\end{description}
\end{frame}

\begin{frame}[fragile]
\frametitle{Structure}
\begin{itemize}
\item
Integrating SAT/SMT solvers into Coq
\begin{enumerate}[1)]
\item Translating formulae into CNF
\item Running the SAT/SMT solver
\item
\begin{enumerate}[a)]
\item
For satisfiable formulae, convert the witness (satisfying assignment) into a proof directly
\item
For unsatisfiable formulae, convert the witness (a resolution proof of $\bot$) into a CIC proof of unsatisfiability
\end{enumerate}
\end{enumerate}
\end{itemize}
\end{frame}

\begin{frame}[fragile]
\frametitle{Resolution}
\begin{itemize}
\item
\begin{Verbatim}[fontsize=\small]
Definition resolution : forall V C D,
    (V \/ C) /\ (~V \/ D) -> C \/ D.
Proof.
    exact (fun V C D H => match H with
    | conj (or_introl v) (or_introl nv) =>
        match nv v with end
    | conj (or_introl v) (or_intror d) => or_intror d
    | conj (or_intror c) _ => or_introl c
    end).
Qed.
\end{Verbatim}
\end{itemize}
\end{frame}

\begin{frame}[fragile]
\frametitle{SMT\_terms.v Form}
\begin{itemize}
\item
\begin{Verbatim}[fontsize=\small]
  Notation fargs := (array _lit) (only parsing).

  Inductive form : Type :=
  | Fatom (_:atom)
  | Ftrue
  | Ffalse
  | Fnot2 (_:int) (_:_lit)
  | Fand (_:fargs)
  | For  (_:fargs)
  | Fimp (_:fargs)
  | Fxor (_:_lit) (_:_lit)
  | Fiff (_:_lit) (_:_lit)
  | Fite (_:_lit) (_:_lit) (_:_lit).
\end{Verbatim}
\end{itemize}
\end{frame}

\begin{frame}[fragile]
\frametitle{SMT\_terms.v Atom}
\begin{itemize}
\item
\begin{Verbatim}[fontsize=\tiny]
  Notation func := int (only parsing).

  Inductive cop : Type := 
   | CO_xH
   | CO_Z0.

  Inductive unop : Type :=
   | UO_xO
   | UO_xI
   | UO_Zpos
   | UO_Zneg
   | UO_Zopp.

  Inductive binop : Type :=
   | BO_Zplus
   | BO_Zminus
   | BO_Zmult
   | BO_Zlt
   | BO_Zle
   | BO_Zge
   | BO_Zgt
   | BO_eq (_ : Typ.type).

  Inductive nop : Type :=
   | NO_distinct (_ : Typ.type).

  Notation hatom := int (only parsing).

  Inductive atom : Type :=
   | Acop (_: cop)
   | Auop (_ : unop) (_:hatom)
   | Abop (_ : binop) (_:hatom) (_:hatom)
   | Anop (_ : nop) (_: list hatom)
   | Aapp (_:func) (_: list hatom).
\end{Verbatim}
\end{itemize}
\end{frame}

\begin{frame}[fragile]
\frametitle{Bvector.reflect}
\begin{itemize}
\item
\begin{Verbatim}[fontsize=\small]
Coq < Require Import Bvector.

Coq < Print Bvector.
Bvector = Vector.t bool
     : nat -> Set

Argument scope is [nat_scope]

Coq < Print reflect.
Inductive reflect (P : Prop) : bool -> Set :=
    | ReflectT : P -> reflect P true
    | ReflectF : ~ P -> reflect P false

For reflect: Argument scopes are [type_scope bool_scope]
For ReflectT: Argument scopes are [type_scope _]
For ReflectF: Argument scopes are [type_scope _]
\end{Verbatim}
\end{itemize}
\end{frame}


\begin{frame}[fragile]
\frametitle{Trace.v's trace}
\begin{itemize}
\item
\begin{Verbatim}[fontsize=\tiny]
Section trace.
  (* We are given a certificate, a checker for it (that modifies a
     state), and a proof that the checker is correct: the state it
     returns must be valid and well-formed. *)
  Variable step : Type.
  Variable check_step : S.t -> step -> S.t.
  Variable rho : Valuation.t.
  Definition _trace_ := Structures.trace step.

  (* A checker for such a trace *)
  Variable is_false : C.t -> bool.
  Hypothesis is_false_correct : forall c, is_false c -> ~ C.interp rho c.
 
  Definition _checker_ (s: S.t) (t: _trace_) (confl:clause_id) : bool :=
    let s' := Structures.trace_fold check_step s t in
    is_false (S.get s' confl).

  (* Proof of its partial correction: if it returns true, then the
     initial state is not valid *)
  Hypothesis valid_check_step :
    forall s, S.valid rho s -> forall c, S.valid rho (check_step s c).

  Lemma _checker__correct :
    forall s, forall t confl, _checker_ s t confl-> ~ (S.valid rho s).
  Proof.
    unfold _checker_.
    intros s t' cid Hf Hv.
    apply (is_false_correct Hf).
    apply S.valid_get.
    apply Structures.trace_fold_ind; auto.
  Qed.
End trace.
\end{Verbatim}
\end{itemize}
\end{frame}

\begin{frame}[fragile]
\frametitle{Tactic Declaration}
\begin{itemize}
\item In \verb|SMTCoq.v|:
\begin{Verbatim}[fontsize=\small]
Declare ML Module "smtcoq_plugin".
\end{Verbatim}
\item \verb|smtcoq_plugin_standard.mlpack| is a list of OCaml modules that get packed into a \verb|.cmxs| archive (OCaml's version of a \verb|.a|)
\item The \verb|Zchaff| and \verb|Verit| modules have a top-level \verb|let tactic () =| declaration which is the entry point for when the tactics get called in Coq.
\end{itemize}
\end{frame}

\begin{frame}[fragile]
\frametitle{zchaff.ml's tactic}
\begin{itemize}
\item
\begin{Verbatim}[fontsize=\tiny]
let core_tactic env sigma t =
  SmtTrace.clear ();
  let (forall_let, concl) = Term.decompose_prod_assum t in
  let a, b = get_arguments concl in
  let reify_atom = Atom.create () in
  let reify_form = Form.create () in
  let (body_cast, body_nocast) =
    if ((Term.eq_constr b (Lazy.force ctrue)) || (Term.eq_constr b (Lazy.force cfalse))) then
      let l = Form.of_coq (Atom.get reify_atom) reify_form a in
      let l' = if (Term.eq_constr b (Lazy.force ctrue)) then Form.neg l else l in
      let atom_tbl = Atom.atom_tbl reify_atom in
      let pform_tbl = Form.pform_tbl reify_form in
      let max_id_confl = make_proof pform_tbl atom_tbl (Environ.push_rel_context forall_let env) reify_form l' in
      build_body reify_atom reify_form (Form.to_coq l) b max_id_confl
    else
      let l1 = Form.of_coq (Atom.get reify_atom) reify_form a in
      let l2 = Form.of_coq (Atom.get reify_atom) reify_form b in
      let l = Form.neg (Form.get reify_form (Fapp(Fiff,[|l1;l2|]))) in
      let atom_tbl = Atom.atom_tbl reify_atom in
      let pform_tbl = Form.pform_tbl reify_form in
      let max_id_confl = make_proof pform_tbl atom_tbl (Environ.push_rel_context forall_let env) reify_form l in
      build_body_eq reify_atom reify_form
    (Form.to_coq l1) (Form.to_coq l2) (Form.to_coq l) max_id_confl
  in
  let compose_lam_assum forall_let body =
    List.fold_left (fun t rd -> Term.mkLambda_or_LetIn rd t) body forall_let in
  let res_cast = compose_lam_assum forall_let body_cast in
  let res_nocast = compose_lam_assum forall_let body_nocast in
  (Structures.tclTHEN
     (Structures.set_evars_tac res_nocast)
     (Structures.vm_cast_no_check res_cast))

let tactic () = Structures.tclTHEN Tactics.intros (Structures.mk_tactic core_tactic)
\end{Verbatim}
\end{itemize}
\end{frame}

\begin{frame}[fragile]
\frametitle{verit.ml's tactic}
\begin{itemize}
\item
\begin{Verbatim}[fontsize=\tiny]
let call_verit rt ro fl root =
  let (filename, outchan) = Filename.open_temp_file "verit_coq" ".smt2" in
  export outchan rt ro fl;
  close_out outchan;
  let logfilename = (Filename.chop_extension filename)^".vtlog" in

  let command = "veriT --proof-prune --proof-merge --proof-with-sharing --cnf-definitional
    --disable-ackermann --input=smtlib2 --proof="^logfilename^" "^filename in
  Format.eprintf "%s@." command;
  let t0 = Sys.time () in
  let exit_code = Sys.command command in
  let t1 = Sys.time () in
  Format.eprintf "Verit = %.5f@." (t1-.t0);
  if exit_code <> 0 then
    failwith ("Verit.call_verit: command "^command^
          " exited with code "^(string_of_int exit_code));
  try
    import_trace logfilename (Some root)
  with
    | VeritSyntax.Sat -> Structures.error "veriT can't prove this"

let tactic () =
  clear_all ();
  let rt = Btype.create () in
  let ro = Op.create () in
  let ra = VeritSyntax.ra in
  let rf = VeritSyntax.rf in
  SmtCommands.tactic call_verit rt ro ra rf

\end{Verbatim}
\end{itemize}
\end{frame}

%\begin{frame}[fragile]
%\frametitle{SMT\_terms.v Atom}
%\begin{itemize}
%\item
%\begin{Verbatim}[fontsize=\small]
%foo
%\end{Verbatim}
%\end{itemize}
%\end{frame}

\begin{frame}[fragile]
\frametitle{Bibleography}
\begin{itemize}
\item ``A Modular Integration of SAT/SMT Solvers to Coq through Proof Witnesses''
\Verb[fontsize=\scriptsize]|http://prosecco.gforge.inria.fr/personal/|
\Verb[fontsize=\scriptsize]|ckeller/Documents-recherche/Publications/cpp11.pdf|
\item ``SMTcoq'' \Verb[fontsize=\scriptsize]|https://github.com/smtcoq/smtcoq|
\item ``The Coq Proof Assistant Reference Manual'' \Verb[fontsize=\scriptsize]|https://github.com/coq/coq/tree/v8.7/doc/refman|
\end{itemize}
\end{frame}

\end{document}
